#  (PART\*) Part one - Soil Legacy Data  {-}

#  Evaluating Soil Legacy Data Sampling for DSM {#legacy_data} 

Modelling techniques in Digital Soil Mapping involve the use of sampling point soil data, with its associated soil properties database, and a number of environmental covariates that will be used to ascertain the relationships of soil properties and the environment to then generalize the findings to locations where no samples have been compiled.

In soil sampling design, a crucial issue is to determine both the locations and the number of the samples to be compiled. In an optimal situation, soil sample database should adequately cover all the environmental diversity space in the study area with a frequency relative to the extent of the diversity in the environmental covariates.

When dealing with legacy soil data, a question that arises is if the data is representative of the environmental diversity within the study area. In this Chapter we present a method to answer this question and to build an alternative how many samples can be retrieved to cover the same environmental space as the existing soil data. The method follows the main findings in [@Malone] and developed as `R` scripts. 

We adapted the original scripts to make use of  vector `'.shp'` and raster `'.tif'` files, as these are data formats commonly used by GIS analysts and in which both soil and environmental data is often stored. We also made some changes in order to simplify the number of R packages and to avoid the use of deprecated packages as it appears in the original code.


##  Data Preparation

We must first load the required packages and data for the analyses. We make use of the packages `sp` and `terra` to manipulate spatial data, `clhs` for Conditioned Latin Hypercube Sampling, `entropy` to compute Kullback-Leibler (KL) divergence indexes, `tripack` for Delaunay triangulation and `manipulate` for interactive plotting within RStudio. Ensure that all these packages are installed in your system before the execution of the script.

```{r load_packages, eval=TRUE, include=F}

# Load packages as a vector objects
  packages <- c("sp", "terra", "clhs", "entropy", "tripack","manipulate","dplyr","plotly") # Create a vector of packages to use
  lapply(packages, require, character.only = TRUE) # Load packages
  rm(packages) # Remove object to save memory space 
  
```

In this manual we recommend to use a common structure for the project, where R scripts appear in the root of the working directory and data files are in a `data/` directory within the root directory (`shp` and `tif` files should be within the sub folders `data/shapes` and `data/rasters` respectively. Following this recommendation simplifies the definition of paths and execution of the scripts. If users want to change their storage paths they have to be properly adjusted in the script.

We define the working directory to the directory in which the actual file is located and load the soil legacy sampling points and the environmental rasters from the `data` folder. To avoid the definition of each environmental covariate, we first retrieve all files with the `.tif` extension and then create a `SpatRaster` object with all of them in a row.


```{r setup_wd, eval=FALSE, include=TRUE}

# Set working directory to source file location
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

```

```{r load_wd, eval=TRUE, include=FALSE}

# Set working directory to source file location
  setwd("/Users/luislado/Dropbox/Github/Sampling_Design_GSP")

```

```{r fig-load_data, eval=TRUE, include=TRUE, fig.cap="Plot of the covariates"}

## Load soil legacy point data
  p.dat <- terra::vect("data/shapes/legacy_soils.shp")

## Load raster covariate data----
  # Read Spatial data covariates as rasters with terra
  rasters <- "data/rasters"
  cov.dat <-  list.files(rasters, pattern = "tif$",  recursive = TRUE, full.names = TRUE)
  cov.dat <- terra::rast(cov.dat)
  # Plot covariates
  plot(cov.dat)
  
```



## Representativeness of the Legacy Soil Data

The next step involves the determination of the distributions of environmental values in the soil samples data and its comparison with the existing distributions of each environmental variable to determine the representativeness of the soil points in the environmental space.

The comparison of distributions is performed through the Kullback-Leibler divergence (KL). It is a measure used to quantify the difference between two probability distributions.
KL-divergence compares an 'objective' or reference probability distribution (here, the distribution of covariates in the complete covariate space - P) with a 'model' or approximate probability distribution (the space of covariates in the soil samples - Q). The main idea is to determine how much information is lost when Q is used to approximate P. In other words, KL-divergence measures how much the Q distribution deviates from the P distribution.

We cross soil and environmental data to create a dataset with the values of the environmental parameters at the locations of the soil samples. 

```{r merge_data, eval=T, include=TRUE}

# Extract environmental data from rasters at soil locations ----
  p.dat_I <- terra::extract(cov.dat, p.dat)
  p.dat_I <- na.omit(p.dat_I) # Remove soil points outside study area
  str(p.dat_I)
  
```

We first calculate a `'n-matrix'` with the values of the covariates dividing their distributions into `'n'` equally-spaced bins. Each bin captures the environmental variability within its interval in the total distribution. In this exercise, `'n'` equals to 25. The result is a  26×4 matrix, where the rows represent the upper and lower limit of the bin and (thus, 26 rows are required to represent 25 bins), and 4 correspond to the number of variables used as environmental proxies.

```{r P_matrix, eval=T, include=TRUE}

# Define Number of bins
  nb<- 25
  #quantile matrix (of the covariate data)
  q.mat<- matrix(NA, nrow=(nb+1), ncol= nlyr(cov.dat))
  j=1
  for (i in 1:nlyr(cov.dat)){ #note the index start here
  #get a quantile matrix together of the covariates
    ran1 <- minmax(cov.dat[[i]])[2] - minmax(cov.dat[[i]])[1]
    step1<- ran1/nb 
    q.mat[,j]<- seq(minmax(cov.dat[[i]])[1], to = minmax(cov.dat[[i]])[2], by =step1)
    j<- j+1}

```

From this matrix, we compute the hypercube matrix of covariates in the whole covariate space.


```{r P_hypercube_matrix, eval=T, include=TRUE}

# Hypercube of "objective" distribution (P) - covariates
  cov.dat.df <- as.data.frame(cov.dat) # convert SpatRaster to dataframe
  cov.mat<- matrix(1, nrow=nb, ncol=ncol(q.mat))
    for (i in 1:nrow(cov.dat.df)){ # the number of pixels
      cntj<- 1 
    for (j in 1:ncol(cov.dat.df)){ #for each column
      dd<- cov.dat.df[i,j]  
      for (k in 1:nb){  #for each quantile
        kl<- q.mat[k, cntj] 
        ku<- q.mat[k+1, cntj] 
        if (is.na(dd)) {
          print('Missing')
        }
        else if (dd >= kl & dd <= ku){cov.mat[k, cntj]<- cov.mat[k, cntj] + 1} 
      }
      cntj<- cntj+1
    }
  }

```

We then calculate the hypercube matrix of covariates in the sample space.


```{r Q_hypercube_matrix, eval=T, include=TRUE}

# Sample data hypercube
  h.mat<- matrix(1, nrow=nb, ncol=ncol(q.mat))
  
  for (ii in 1:nrow(p.dat_I)){ # the number of observations
    cntj<- 1 
    for (jj in 2:ncol(p.dat_I)){ #for each column
      dd<- p.dat_I[ii,jj]  
      for (kk in 1:nb){  #for each bin
        kl<- q.mat[kk, cntj] 
        ku<- q.mat[kk+1, cntj] 
        if (dd >= kl & dd <= ku){h.mat[kk, cntj]<- h.mat[kk, cntj] + 1}
      }
      cntj<- cntj+1
    }
  }

```

* **KL-divergence**

We calculate the KL-divergence to measure how much the distribution of covariates in tbe sample space (Q) deviates from the distribution of covariates in the complete study area space (P).
  
```{r KL_divergence, eval=T, include=TRUE}
## Compare covariate distributions in P and Q with Kullback-Leibler (KL) divergence
    kl.index <-c()
    for(i in 1:ncol(cov.dat.df)){
      kl <-    KL.empirical(c(cov.mat[,i]), c(h.mat[,i]))
      kl.index <- c(kl.index,kl)
      klo <-  mean(kl.index)
    }
    print(kl.index) # KL divergences of each covariate
    print(klo) # KL divergence in the existing soil samples

```

The KL-divergence is always greater than or equal to zero, and reaches its minimum value (zero) only when P and Q are identical. Thus, lower values of KL-divergence are indicative of a good match between both the sample and the study area spaces, indicating that the sample space is a fair representation of the environmental conditions in the study area.

In this case, the KL-divergence value is `r round(klo,3)`, indicating that the legacy samples capture most of the environmental variability in the study area.

* **Percent of representativeness in relation to the overall environmental conditions**

Finally, we can also determine the degree in which our legacy soil dataset is representative of the existing environmental conditions in the study area. For that, we calculate the proportion of pixels in the study area that would fall within the convex hull polygon delineated upon the environmental conditions found at the soil legacy data locations only. The convex hull polygon is created upon a Principal Component transformation of the covariate data in the soil legacy data and using the outter limits of the scores of the points projected on the two main Components.


```{r fig-legacy_data_adequacy, eval=T, include=TRUE}
#### Representativeness of the Legacy Dataset: ----
  ## Calculate the proportion of "env. variables" in the covariate spectra that fall within the convex hull of variables in the "environmental sample space"
  
  # Principal component of the legacy data sample
    pca.s = prcomp(p.dat_I[,2:(ncol(cov.dat.df)+1)],scale=TRUE, center=TRUE)
    scores_pca1 = as.data.frame(pca.s$x)
  # Plot the first 2 principal components and convex hull
    rand.tr <- tri.mesh(scores_pca1[,1],scores_pca1[,2],"remove") # Delaunay triangulation 
    rand.ch<-convex.hull(rand.tr, plot.it=F) # convex hull
    pr_poly = cbind(x=c(rand.ch$x),y=c(rand.ch$y)) # save the convex hull vertices
    plot(scores_pca1[,1], scores_pca1[,2], xlab="PCA 1", ylab="PCA 2", xlim=c(min(scores_pca1[,1:2]), max(scores_pca1[,1:2])),ylim=c(min(scores_pca1[,1:2]), max(scores_pca1[,1:2])), main='Convex hull of soil legacy data')
    lines(c(rand.ch$x,rand.ch$x[1]), c(rand.ch$y,rand.ch$y[1]),col="red",lwd=1) # draw the convex hull (domain of legacy data)

  # PCA projection of study area population onto the principal components
    PCA_projection<- predict(pca.s, cov.dat.df) # Project study area population onto sample PC
    newScores = cbind(x=PCA_projection[,1],y=PCA_projection[,2]) # PC scores of projected population
  
  # Plot the polygon and all points to be checked
    plot(newScores, xlab="PCA 1", ylab="PCA 2", xlim=c(min(newScores[,1:2]), max(newScores[,1:2])), ylim=c(min(newScores[,1:2]), max(newScores[,1:2])), col='black', main='Environmental space plots over the convex hull of soil legacy data')
    polygon(pr_poly,col='#99999990')
  # Check which points fall within the polygon
    pip <- point.in.polygon(newScores[,2], newScores[,1], pr_poly[,2],pr_poly[,1],mode.checked=FALSE)
    newScores <- data.frame(cbind(newScores, pip))
  # Plot points outside convex hull  
    points(newScores[which(newScores$pip==0),1:2],pch='X', col='red')
  # Proportion of the conditions in the study area that fall within the convex hull
    sum(newScores$pip)/nrow(newScores)*100 

```

This indicates that `r round(sum(newScores$pip)/nrow(newScores)*100,1)`\% of the existing conditions in the study area fall within the convex hull delineated with the data in the  soil samples, showing the adequacy of the proposed legacy data for DSM.
