
# Stratified Sampling Design

Stratified random sampling is a technique where the study area is divided into different `groups` or `strata` based on certain environmental traits and a number of random samples are taken from within each group. One of the primary advantages of stratified sampling is its ability to capture the diversity within a population by making sure each group is represented. It can provide a more accurate reflection of the entire population compared to random sampling, especially when the groups are distinct and have unique qualities. This approach is particularly beneficial when certain subgroups within the population are specifically noteworthy. It also allows for more precise estimates with a smaller total sample size compared to simple random choice. Stratified sampling presents some disadvantages. Achieving effective categories requires a proper definition and delineation of the initial information to create the `strata`. The classification of the environmental information into categories and ensuring fair portrayal of each can be intricate and time–taking and mislabeling elements into an improper group can lead to skewed outcomes.

## General Procedure

The creation of a  stratified random sampling design involves the identification of relevant features describing the environmental diversity in the area (soil and land use are the environmental variables generally used to define strata), delineation of the strata, determination of the number of samples to distribute to each stratum, followed by random sampling within it. By identifying relevant classes, combining them to define strata and allocating an appropriate number of samples to each stratum, a representative sample can be obtained. Random sampling within each stratum helps to ensure that the sample is unbiased and provides a fair representation of the overall conditions in the area.

The first question is about how many samples must be retrieved from each strata. The sampling scheme starts with the definition of the total number of samples to collect. In this case, the determination of the sample size is a complex and highly variable process based, among others, on the specific goals of the study, the variability of environmental proxies, the statistical requirements for accuracy and confidence, as well as additional considerations such as accessibility, costs and available resources. The optimal number of samples can be determined following the method proposed in Chapter 2 of this manual. The number of samples within each stratum is calculated using an area–weighted approach taking into account the relative area of each stratum. The sampling design in this section must also comply with the following requirements:

   *  All sampling strata must have a minimum size of 100 hectares.
   *  All sampling strata must be represented by at least 2 samples.

This sampling process ensures the representativeness of the environmental combinations present across the area while maintaining an efficient and feasible field sampling campaign.

### Strata creation

We must determine the kind of information that will be used to construct the `strata`. In this manual, we present a simple procedure to build strata based on data from two environmental layers: soil groups and land use classification data. The information should be provided in the form of vector shapefiles with associated information databases. The data on both sets often comprises a large number of categories, that would lead to a very large number of `strata`. Thus, it is desirable to make an effort of aggregating similar categories within each input data set, to reduce, as much as possible, the number of categories while still capturing the most of the valuable variability in the area.

The fist step is to set–up the RStudio environment and load the required packages:


```{r strata–setup, eval=TRUE, include=FALSE, warning=FALSE}

# Load packages as a vector objects
  packages <- c("sf", "terra", "tidyverse", "rmapshaper", "units","plyr", "mapview", "leaflet","stars","sgsR")
  lapply(packages, require, character.only = TRUE) # Load packages
  rm(packages) # Remove object to save memory space 

  # Set working directory to source file location
  #setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

```


```{r load_wd_strata, eval=TRUE, include=FALSE, warning=FALSE}

  ## Set working directory to source file location
  #setwd("/Users/luislado/Dropbox/Github/Sampling_Design_GSP")

```


We must define the number of samples to distribute in the sampling design and the soil and land use information layers to build the strata. We also define a REPLACEMENT parameter to account for a reduction of the sampling area according to a certain area using predefined bounding–box, that can be also here defined.


```{r strata–parameters, eval=TRUE, include=FALSE, echo=FALSE}

  # Define number of samples
    n <- 242

  # Load soil data
    soil <- st_read(paste0(shp.path,"/soils.shp"),promote_to_multi = FALSE)  
  # Load Landcover data
    lc <- st_read(paste0(shp.path,"/landcover.shp"))
    
    # Define a bounding box (optional). If REPLACEMENT = TRUE, then the bounding–box shapefile (bb), must be specified and the uncomment the line in the code below
    BOUNDING = FALSE
    # bb <- st_read("your_bounding_box.shp")
    REPLACEMENT = FALSE
    distance.buffer = 500 # Buffer distance for samples

```

We proceed with the calculation of soil groups. In this example, soil information is stored in the field `SRG`. We have analysed the extent to which the information in this field can be synthesized to eliminate redundancy when creating the `strata`. ^[This exploratory work is a prerequisite and must be adapted specifically to each soil and land use dataset].
The results are shown in \@ref(fig:fig-10)

```{r soil–aggregation, eval=TRUE, include=FALSE, warning=FALSE}
  
  # Clip soil data  by bounding area if defined
  if (BOUNDING) {
    soil <- st_intersection(soil, bb)
  } else {
    soil <- soil
  }
    # Check geometry
    soil <- st_make_valid(soil)
    soil <- st_cast(soil, "MULTIPOLYGON")
    #Aggregate information by reference soil group (SRG)
      #unique(soil$RSG)
    # Remove NAs (water bodies are deleted)
    soil <- soil[!is.na(soil$RSG),]
      #unique(soil$RSG)
    # Dissolve polygons by reference soil group
    soil <- soil %>% group_by(RSG) %>% dplyr::summarize() 
 
```



```{r fig-10, fig.cap="Plot of the soil classes", eval=TRUE, include=TRUE}
  
  ## Plot aggregated soil classes
    map = leaflet(leafletOptions(minZoom = 8)) %>%
      addTiles()
    mv <- mapview(soil["RSG"], alpha=0, homebutton=T, layer.name = "soils", map=map)
    mv@map
    #ggplot() + geom_sf(data=soil, aes(fill = factor(RSG)))

```


A similar procedure is performed on the land use dataset.


```{r landuse–aggregation, eval=TRUE, include=FALSE, warning=FALSE}
  # Calculate Land use Groups

  # Clip land use data  by bounding area if exist
  if (BOUNDING) {
    lc <- st_intersection(lc, bb)
  } else {
    lc <- lc
  }
  # Check geometry
  lc <- st_make_valid(lc)
  # Agregate units
  #unique(lc$landcover)
  lc$landcover[lc$landcover=="Broadleaf forest"]<- "Forest"
  lc$landcover[lc$landcover=="Needleleaf forest"]<- "Forest"
  lc$landcover[lc$landcover=="Mixed forest"]<- "Forest"
  lc <- lc[lc$landcover!="Water body",]
  #unique(lc$landcover)
  
  # Dissolve polygons by reference soil group
  lc <- lc %>% group_by(landcover) %>% dplyr::summarize() 
    
 
```

Figure \@ref(fig:fig-11) shows the landcover classes to build the strata.

```{r fig-11, fig.cap="Plot of the land use classes", eval=TRUE, include=TRUE}

# Plot map with the land cover information
    map = leaflet() %>%
      addTiles()
    mv <- mapview(lc["landcover"], alpha=0, homebutton=T, layer.name = "Landcover", map=map)
    mv@map
    #ggplot() + geom_sf(data=lc, aes(fill = factor(landcover)))
```

To create the soil–land use `strata` we must combine both classified datasets.


```{r combine–data, eval=TRUE, include=TRUE, warning=FALSE} 
  # Combine soil and land use layers
  sf_use_s2(FALSE)
  soil_lc <- st_intersection(st_make_valid(soil), st_make_valid(lc))  
  soil_lc$soil_lc <- paste0(soil_lc$RSG, "_", soil_lc$landcover)
  soil_lc <- soil_lc %>% dplyr::select(soil_lc, geometry)
  
```

Finally, to comply with the initial requirements of the sampling design, we calculate the areas of each polygon, delete all features with extent lesser than 100 has.  

```{r select–area, eval=TRUE, include=FALSE} 
  
  # Select by Area. Convert to area to ha and select polygons with more than 100 has
  soil_lc$area <- st_area(soil_lc)/10000 
  soil_lc$area <- as.vector(soil_lc$area)
  soil_lc <- soil_lc %>% 
    group_by(soil_lc) %>% 
    mutate(area = sum(area))
  soil_lc <- soil_lc[soil_lc$area > 100,]
  plot(soil_lc[1])
  
  # Replace blank spaces with underscore symbol to keep names uniform
  soil_lc$soil_lc <- str_replace_all(soil_lc$soil_lc, " ", "_")
  
  # Create a column of strata numeric codes
  soil_lc$code <- as.character(as.numeric(as.factor(soil_lc$soil_lc)))
  
  # List final strata
  unique(soil_lc$soil_lc)

  # Write final sampling strata map
  #st_write(soil_lc, paste0(results.path,"strata.shp"), delete_dsn = TRUE)

```

The final strata map is shown in Figure  \@ref(fig:fig-12).

```{r fig-12, eval=TRUE, include=TRUE, fig.cap="Plot of strata"}

   # Plot final map of stratum
  map = leaflet(options = leafletOptions(minZoom = 8.3)) %>%
  addTiles()
  mv <- mapview(soil_lc["soil_lc"], alpha=0, homebutton=T, layer.name = "Strata", map=map)
  mv@map
  
```


## Stratified random sampling

This example demonstrates how to establish a stratified random sampling approach within the previously defined strata polygons. The allocation of sample points is proportionate to the stratum areas, with the condition that each stratum must contain a minimum of 2 samples. The determination of sampling points, referred to as `'target points'`, is made during the initial phase of the sampling design and takes into consideration factors such as the area to be sampled, budget constraints and available personnel. Additionally, a set number of `'replacement points'` must be designated to act as substitutes for 'target points' in cases where some of the original target points cannot be accessed or sampled. These 'replacement points' are systematically indexed, with each index indicating which 'target point' it serves as a substitute for.


```{r random–sampling, eval=TRUE, include=FALSE}
  ## Stratified Simple Random Sampling over Soil/Land use Strata
  
  # Read strata shapefile
  polygons <- st_read(paste0(results.path,"strata.shp"))
  if(REPLACEMENT){
    polygons = st_intersection(polygons,distance.buffer)
  }
  

    polygons$area <- st_area(polygons) # calculate the area of each polygon
    
    # Create a new column to group polygons by a common attribute
    polygons$group <- polygons$soil_lc
    # Drop units to allow computations
    polygons <- drop_units(polygons)
    
    # Calculate the total area of all polygons in each group
    group_areas <- polygons %>%
      dplyr::group_by(group)  %>% 
      dplyr::summarize(total_area = sum(area))
    # Add a code to each group
    group_codes <- polygons %>% group_by(group) %>%
      dplyr::summarize(code = first(code)) 
    
    group_areas <- left_join(group_areas,st_drop_geometry(group_codes), by = "group")
    
    # Ensure minimum of 2 samples at each polygon in each group
    group_areas$sample_count <- 2
    
    # Calculate the number of samples per group based on relative area
    group_areas$sample_count <- group_areas$sample_count+round(group_areas$total_area/sum(group_areas$total_area) * 
                                                                 (n-sum(group_areas$sample_count)))
    
    while (sum(group_areas$sample_count) != n) {
      if (sum(group_areas$sample_count) > n) {
        # Reduce sample count for the largest polygon until total count is n
        max_index <- which.max(group_areas$sample_count)
        group_areas$sample_count[max_index] <- group_areas$sample_count[max_index] - 1
      } else {
        # Increase sample count for the smallest polygon until total count is n
        min_index <- which.min(group_areas$sample_count)
        group_areas$sample_count[min_index] <- group_areas$sample_count[min_index] + 1
      }
    }
    #sum(group_areas$sample_count) 
    
    polygons <- left_join(polygons, st_drop_geometry(group_areas), by = c("soil_lc"="group"))
    polygons <- dplyr::select(polygons, soil_lc, code.x, sample_count, geometry)
    
    # Generate random points within each strata of size 3 times the required samples for each strata ----
    x <- spatSample(x = vect(group_areas), size = group_areas$sample_count * 3, method = "random")
    
    # Compute sampling points for strata
    z <- x %>% 
      st_as_sf() %>% 
      dplyr::group_by(code) %>% 
      dplyr::mutate(sample_count = as.numeric(sample_count),
             order = seq_along(code),
             ID = paste0(code, ".", order),
             type = ifelse(sample_count >= order, "Target", "Replacement")) %>% 
      vect()
    
    # Find missing samples
    missing.data <- left_join(group_areas,data.frame(z) %>%
                                dplyr::filter(type=="Target") %>%
                                dplyr::group_by(code) %>%
                                tally()) %>%
      dplyr::mutate(diff=sample_count-n)
    
    # Determine missing sampled strata
    missing.strata <- which(is.na(missing.data$diff))
    
    # Determine missing sampling point in strata (undersampled strata)
    missing.sample = which(missing.data$diff != 0)
    missing.number <- as.numeric(unlist(st_drop_geometry(missing.data[(missing.sample <- which(missing.data$diff != 0)),7])))
    
    # Compute sampling points for missing sampled strata
    x.missing.strata <- x[1]
    x.missing.strata$sample_count<- 0
    
    for(i in missing.strata){
      xx.missing.strata <- x[1]
      xx.missing.strata$sample_count<- 0
      nn=0
      while (sum(xx.missing.strata$sample_count) < 
             group_areas[group_areas$code==i,][["sample_count"]]*5) {
        
        while(nn < group_areas[group_areas$code==i,][["sample_count"]]*3){
          my.missing.strata <- spatSample(x = vect(group_areas[group_areas$code %in% i,]),
                                          size =  group_areas[group_areas$code==i,][["sample_count"]]*5,
                                          method = "random")
          nn <- nn + nrow(data.frame(my.missing.strata))
        }
        xx.missing.strata <- rbind(xx.missing.strata,my.missing.strata)
        print(sum(xx.missing.strata$sample_count))
      }
      print(i)
      print(xx.missing.strata)
      x.missing.strata <- rbind(x.missing.strata,xx.missing.strata)
    }
    
    # Join initial sampling with missing sampling strata data
    x <- rbind(x, x.missing.strata)
    
    # Compute sampling points for missing samples (random sampling)
    x.missing.sample <- x[1]
    
    for(i in missing.sample){
      xx.missing.sample <- x[1]
      xx.missing.sample$sample_count<- 0
      while (sum(xx.missing.sample$sample_count) < (group_areas[group_areas$code==i,][["sample_count"]]*3)) {
        my.missing.sample <- spatSample(x = vect(group_areas[group_areas$code %in% i,]),
                                        size = as.numeric(vect(group_areas[group_areas$code %in% i,])[[4]])+
                                          (group_areas[group_areas$code==i,][["sample_count"]]*3), method = "random")
        
        xx.missing.sample <- rbind(xx.missing.sample,my.missing.sample)
        print(sum(xx.missing.sample$sample_count))
      }
      print(i)
      print(xx.missing.sample)
      x.missing.sample <- rbind(x.missing.sample,xx.missing.sample)
    }
    
    # Join initial sampling with missing sampling strata data and with missing samples 
    x <- rbind(x, x.missing.sample)
    
    # Remove extra artificial replacements 
    x <- x[x$sample_count > 0,]
    
    # Convert to Shapefile
    z <- x %>% 
      st_as_sf() %>% 
      dplyr::group_by(code) %>% 
      dplyr::mutate(sample_count = as.numeric(sample_count),
             order = seq_along(code),
             ID = paste0(code, ".", order),
             type = ifelse(sample_count >= order, "Target", "Replacement")) %>% 
      vect()
    
    # Export data to Shapefile ----
      
      # Write sampling points to shp
      #writeVector(z, paste0(results.path,"strat_randm_samples.shp"), overwrite=TRUE)
    
      # Check whether the number of initial target points equals the final target points 
      #n;nrow(z[z$type=="Target",])

```


Results are shown in Figure \@ref(fig:fig-13).

```{r fig-13, fig.cap="Plot of strata and random target and replacement points", eval=TRUE, include=TRUE}

  map = leaflet(options = leafletOptions(minZoom = 8.3)) %>%
        addTiles()
  mv <- mapview(soil_lc["soil_lc"], alpha=0, homebutton=T, layer.name = "Strata") + 
        mapview(sf::st_as_sf(z), zcol = 'type', color = "white", col.regions = c('royalblue', 'tomato'), cex=3, legend = TRUE,layer.name = "Samples")
  mv@map

```




## Stratified random sampling for large areas

The implementation of a stratified random sampling, along with target and replacement points, can present operating difficulties when dealing with areas of significant size and with locations that are hard to reach. To address this issue, the sampling approach can be modified by excluding areas with limited accessibility.

This modification can streamline fieldwork operations and establish a feasible sampling method while still retaining the essence of the stratified random sampling framework. By excluding areas with limited accessibility, the sampling design can be adjusted to ensure a more practical and effective approach to data collection.

 * **Delineation of sampling accessibility:** The sampling area can be further limited based on accessibility considerations. Areas with very limited accessibility, defined as regions located more than 1 kilometre away from a main road or access path, may be excluded from sampling areas. To accomplish this, a map of main roads and paths can be used to establish a sampling buffer that includes areas within a 1–kilometre buffer around the road infrastructures. This exclusion helps to eliminate the most remote and challenging–to–access areas. An additional layer of accessibility information can be incorporated based on population distribution in the country, considering that, if population is present, there is a high change that points in the surroundings can be accessible for sampling. In this case, populated nuclei are vectorized into points and a 250–meter buffer is then generated around each point. These resulting areas can be then added to the 1–kilometre buffer around the roads, which collectively defined the final sampling area.
 
 * **Substitution of replacement points with replacement areas in close proximity to the target points:** The sampling design presented before included designated replacement points to serve as substitutes for each target point in the case that it would be inaccessible during fieldwork. However, this approach presented challenges, particularly for large areas, as the replacement point could be located far from the target point, resulting in significant logistical efforts. This limitation posed a risk of delays in completing the sampling campaign within the allocated time frame. To address this challenge, an alternative strategy is to replace the idea of replacement points with replacement areas situated in the immediate vicinity of the target point. The replacement area for each target point is now confined within a 500–meter buffer surrounding the target and falls within the same sampling stratum. This approach concentrates sampling and replacement activities within a specific geographic area, streamlining the overall process. By reducing the need for extensive travel, this method enhances efficiency and facilitates sample collection. Figure 2 illustrates the distribution of sampling points and replacement areas for visualization.
 
  * **Additional area exclusion:** Some areas can be identified as not suitable for sampling purposes. This is the case of certain natural protected areas, conflict regions presenting risks for field operators, etc. These areas must be identified masked at an initial stage of the design to exclude them from the sampling strata.
 
 The procedure is the same as that previously presented, with the difference that buffers and exclusion areas must be masked–out from the strata map before performing the random sampling.
 
 
```{r replacement–areas, eval=FALSE, include=TRUE} 

 # Compute sampling areas WITH REPLACEMENT -----
  if(REPLACEMENT){
      # Load strata
      soil_lc <- st_read(paste0(results.path,"strata.shp"))

      # Read sampling. points from previous step
      z <- st_read(paste0(results.path,"sampling_points.shp"))
      
      # Define buffer of 500 meters (coordinate system must be in metric base)
      buf.samples <- st_buffer(z, dist=distance.buffer)
    
      # Intersect buffers
      samples_buffer = st_intersection(soil_lc, buf.samples)
      samples_buffer <- samples_buffer[samples_buffer$type=="Target",]
      samples_buffer <- samples_buffer[samples_buffer$soil_lc==samples_buffer$group,]
      # Save Sampling areas
      #st_write(samples_buffer, paste0('../soil_sampling/JAM/replacement_areas_', samples.buffer, '.shp'), delete_dsn = TRUE)
      
      # Write target points only
      targets <- z[z$type=="Target",]
      #st_write(targets, '../soil_sampling/JAM/sampling_points_TAR.shp', delete_dsn = TRUE)
  }

```


## Stratified regular sampling

The procedure for creating a stratified regular sampling design is identical to that presented for stratified random sampling, with the only distinction that the locations of the sampling points are distributed in a regular spatial grid. This transformation is achieved by changing the method from 'random' to 'regular' in the spatSample functions within the script above. 


```{r regular–sampling, eval=TRUE, include=FALSE}

# Stratified Regular Sampling over Soil/Land use Strata ----
  # Read strata shapefile
  polygons <- st_read(paste0(results.path,"strata.shp"))
  if(REPLACEMENT){
    polygons = st_intersection(polygons,distance.buffer)
  }
  

polygons$area <- st_area(polygons) # calculate the area of each polygon

# Create a new column to group polygons by a common attribute
polygons$group <- polygons$soil_lc
# Drop units to allow computations
polygons <- drop_units(polygons)

# Calculate the total area of all polygons in each group
group_areas <- polygons %>%
  dplyr::group_by(group)  %>% 
  dplyr::summarize(total_area = sum(area))
# Add a code to each group
group_codes <- polygons %>% group_by(group) %>%
  dplyr::summarize(code = first(code)) 

group_areas <- left_join(group_areas,st_drop_geometry(group_codes), by = "group")

# Ensure minimum of 2 samples at each polygon in each group
group_areas$sample_count <- 2

# Calculate the number of samples per group based on relative area
group_areas$sample_count <- group_areas$sample_count+round(group_areas$total_area/sum(group_areas$total_area) * 
                                                             (n-sum(group_areas$sample_count)))

while (sum(group_areas$sample_count) != n) {
  if (sum(group_areas$sample_count) > n) {
    # Reduce sample count for the largest polygon until total count is n
    max_index <- which.max(group_areas$sample_count)
    group_areas$sample_count[max_index] <- group_areas$sample_count[max_index] - 1
  } else {
    # Increase sample count for the smallest polygon until total count is n
    min_index <- which.min(group_areas$sample_count)
    group_areas$sample_count[min_index] <- group_areas$sample_count[min_index] + 1
  }
}
#sum(group_areas$sample_count) 

polygons <- left_join(polygons, st_drop_geometry(group_areas), by = c("soil_lc"="group"))
polygons <- dplyr::select(polygons, soil_lc, code.x, sample_count, geometry)

 
# Generate regular points within each strata of size 3 times the required samples for each strata ----
x <- spatSample(x = vect(group_areas), size = group_areas$sample_count * 3, method = "regular")

# Compute sampling points for strata
z <- x %>% 
  st_as_sf() %>% 
  dplyr::group_by(code) %>% 
  dplyr::mutate(sample_count = as.numeric(sample_count),
         order = seq_along(code),
         ID = paste0(code, ".", order),
         type = ifelse(sample_count >= order, "Target", "Replacement")) %>% 
  vect()

# Find missing samples
missing.data <- left_join(group_areas,data.frame(z) %>%
                            dplyr::filter(type=="Target") %>%
                            dplyr::group_by(code) %>%
                            tally()) %>%
  dplyr::mutate(diff=sample_count-n)

# Determine missing sampled strata
missing.strata <- which(is.na(missing.data$diff))

# Determine missing sampling point in strata (undersampled strata)
missing.sample = which(missing.data$diff != 0)
missing.number <- as.numeric(unlist(st_drop_geometry(missing.data[(missing.sample <- which(missing.data$diff != 0)),7])))

# Compute sampling points for missing sampled strata
x.missing.strata <- x[1]
x.missing.strata$sample_count<- 0

for(i in missing.strata){
  xx.missing.strata <- x[1]
  xx.missing.strata$sample_count<- 0
  nn=0
  while (sum(xx.missing.strata$sample_count) < 
         group_areas[group_areas$code==i,][["sample_count"]]*5) {
    
    while(nn < group_areas[group_areas$code==i,][["sample_count"]]*3){
      my.missing.strata <- spatSample(x = vect(group_areas[group_areas$code %in% i,]),
                                      size =  group_areas[group_areas$code==i,][["sample_count"]]*5,
                                      method = "random")
      nn <- nn + nrow(data.frame(my.missing.strata))
    }
    xx.missing.strata <- rbind(xx.missing.strata,my.missing.strata)
    print(sum(xx.missing.strata$sample_count))
  }
  print(i)
  print(xx.missing.strata)
  x.missing.strata <- rbind(x.missing.strata,xx.missing.strata)
}

# Join initial sampling with missing sampling strata data
x <- rbind(x, x.missing.strata)

# Compute sampling points for missing samples (regular sampling)
x.missing.sample <- x[1]

for(i in missing.sample){
  xx.missing.sample <- x[1]
  xx.missing.sample$sample_count<- 0
  while (sum(xx.missing.sample$sample_count) < (group_areas[group_areas$code==i,][["sample_count"]]*3)) {
    my.missing.sample <- spatSample(x = vect(group_areas[group_areas$code %in% i,]),
                                    size = as.numeric(vect(group_areas[group_areas$code %in% i,])[[4]])+
                                      (group_areas[group_areas$code==i,][["sample_count"]]*3), method = "regular")
    
    xx.missing.sample <- rbind(xx.missing.sample,my.missing.sample)
    print(sum(xx.missing.sample$sample_count))
  }
  print(i)
  print(xx.missing.sample)
  x.missing.sample <- rbind(x.missing.sample,xx.missing.sample)
}

# Join initial sampling with missing sampling strata data and with missing samples 
x <- rbind(x, x.missing.sample)

# Remove extra artificial replacements 
x <- x[x$sample_count > 0,]

# Convert to Shapefile
z <- x %>% 
  st_as_sf() %>% 
  dplyr::group_by(code) %>% 
  dplyr::mutate(sample_count = as.numeric(sample_count),
         order = seq_along(code),
         ID = paste0(code, ".", order),
         type = ifelse(sample_count >= order, "Target", "Replacement")) %>% 
  vect()

# Export data to Shapefile ----
  
  # Write sampling points to shp
  #writeVector(z, paste0(results.path,"sampling_points.shp"), overwrite=TRUE)

  # Check whether the number of initial target points equals the final target points 
  n;nrow(z[z$type=="Target",])

```


```{r fig-14, fig.cap="Plot of strata and regular sampling points", eval=TRUE, include=TRUE}


  map = leaflet(options = leafletOptions(minZoom = 11.4)) %>%
        addTiles()
  mv <- mapview(soil_lc["soil_lc"], alpha=0, homebutton=T, layer.name = "Strata") + 
        mapview(sf::st_as_sf(z), zcol = 'type', color = "white", col.regions = c('royalblue', 'tomato'), cex=3, legend = TRUE,layer.name = "Samples")
  mv@map
  
  
```


## Random Sampling based on a stratified raster

Finally, it is also possible to create a stratified area weighted random sampling using raster strata. The procedure involves the creation of the strata as a raster file and implement a random sampling using the frequencies of the strata as a guideline for distribution of the samples proportionally to their frequencies. This method is easily implemented using the package 'sgsR' [@sgsr]. 

```{r sgsR-random,eval=TRUE, include=TRUE, message=FALSE, warning=FALSE}

 strata <- st_read(paste0(results.path,"strata.shp"),quiet = TRUE)
 strata$code <- as.integer(strata$code)
 
 # Create stratification raster 
  strata <- rast(st_rasterize(strata["code"],st_as_stars(st_bbox(strata), nx = 250, ny = 250)))
  names(strata) <- "strata"
  strata <- crop(strata, nghe, mask=TRUE)
  
  # Create stratified random sampling
  target <- sample_strat(
    sraster = strata,
    nSamp = 200
    )
  target$type <- "target"
  

```

Figure  \@ref(fig:fig-14b)  show the histogram of frequencies of samples over the strata categories respectively.


```{r fig-14b, fig.cap="Frequencies of strata and random target samples", eval=TRUE, include=TRUE, message=FALSE, warning=FALSE}
  # Histogram of frequencies
  calculate_representation(
    sraster = strata,
    existing = target,
    drop=0,
    plot = TRUE 
  )

  # Add index by strata
  target <- target %>% 
    st_as_sf() %>% 
    dplyr::group_by(strata) %>% 
    dplyr::mutate(sample_count = sum(n),
                  order = seq_along(strata),
                  ID = paste0(strata, ".", order)) %>% 
    vect()
  

```


The construction of replacement points is straightforward by creating a new stratified set of samples on the same strata. In this case, the sample size has been incremented 3 times to create 3 replacement points for each target point. 

```{r replacements-raster, eval=TRUE, include=TRUE, message=FALSE, warning=FALSE}
   replacement <- sample_strat(
    sraster = strata,
    nSamp = 200*3
    )
   replacement$type <- "replacement"


```


```{r fig-14c, fig.cap="Frequencies of strata and random replacement samples", eval=TRUE, include=TRUE, message=FALSE, warning=FALSE}
  # Histogram of frequencies
  calculate_representation(
    sraster = strata,
    existing = replacement,
    drop=0,
    plot = TRUE 
  )

     # Add index by strata
  replacement <- replacement %>% 
    st_as_sf() %>% 
    dplyr::group_by(strata) %>% 
    dplyr::mutate(sample_count = sum(n),
                  order = seq_along(strata),
                  ID = paste0(strata, ".", order)) %>% 
    vect()

```

Figures  \@ref(fig:fig-14c)  and \@ref(fig:fig-14d) show the frequencies of replacement samples over the strata categories and the distribution of target and replacement samples respectively.

```{r fig-14d, fig.cap="Plot of raster strata and sampling points", eval=TRUE, include=TRUE}
  # Plot samples over strata
  # plot(strata, main="Strata and random samples")
  # plot(nghe[1], col="transparent", add=TRUE)
  # points(target,col="red")
  # points(replacement,col="royalblue")
  # legend("topleft", legend = c("target","replacement"), pch = 20, xpd=NA, bg="white", col=c("red","royalblue"))
  
  map = leaflet(options = leafletOptions(minZoom = 8.3)) %>%
        addTiles()
  mv <- mapview(soil_lc["soil_lc"], alpha=0, homebutton=T, layer.name = "Strata") + 
        mapview(target, zcol = 'type', color = "white", col.regions = c('tomato'), cex=3, legend = TRUE,layer.name = "Target") + 
        mapview(replacement, zcol = 'type', color = "white", col.regions = c('royalblue'), cex=3, legend = TRUE,layer.name = "Replacement")
  mv@map
  
```

The resulting target and replacement points can finally be stored as shapefiles.

```{r write-raster-pts, eval=FALSE, include=TRUE}
  ## Export to shapefile
   writeVector(target,
                paste0(results.path,"target_pts_raster.shp"), overwrite=TRUE)
   writeVector(replacement,
                paste0(results.path,"replacement_pts_raster.shp"), overwrite=TRUE)

```   
